\chapter{Summary and Outlook}

In this thesis, the importance of relativistic effects in autoionization
processes, especially \ac{ICD}-like
processes, and cluster environments have been discussed.
For this purpose, asymptotic expressions for the relativistic decay width of the ICD
and both, relativistic and non-relativistic asymptotic expressions for the ETMD3
have been derived. Additionally, the non-relativistically well-established
FanoADC-Stieltjes approach
using a partitioning by population has been implemented
in the relativistic program package Dirac,
which allows for the
description of decay width including relativistic effects.
In order to simulate the experimental secondary electron spectra of noble gas
clusters, the model of pairs and triples was introduced and automatized in the
program HARDRoC. It enables the estimation of
decay widths of large systems like clusters from data of the compounds.

In the studies of the atomic Auger process, scalar-relativistic effects were
found to increase the decay width compared to the non-relativistic results
due to larger orbital overlaps of the initial and final states. Especially
in ETMD processes, whose decay widths are governed by the orbital overlap
of the two units involved in the electron transfer, similar significant
decay width influences might be observed.

Throughout all systems containing heavy elements, the spin-orbit coupling
shows a pronounced effect on the secondary electron spectrum by increasing the
number of possible channels and hence, the number of peaks. This feature cannot
be explained using a non-relativistic methodology.
If the decay channels are close to threshold, this energetic splitting
can for a non-relativistically closed decay channel cause some of the
corresponding decay channels in the relativistic description to be open. On
the other hand, not all relativistic channels corresponding to one specific
open non-relativistic channel need to be energetically accessible.

Additionally, geometry has a great impact on the opening and closing of channels
in \ac{ICD}-like processes. The closer the ionized atoms of the final state are,
the lower is the kinetic energy of the secondary electron and at some
internuclear distance the decay channel closes. For the ICD, it is a channel
closing internuclear distance and for the ETMD3 considering a fixed electron
transfer distance, it is a curve along the intersection of the initial and final
state hypersurfaces.

In larger clusters, the additional effect of charge stabilization
shifts the kinetic energy of the secondary
electron. These shifts were treated by using experimentally obtained
ionization energies for exactly the same experimental conditions as for the
secondary electron spectra and by shifting the atomic ionization potentials
by experimentally obtained shifts in homonuclear clusters.
A more predictive way would be to simulate the influence of the surrounding
atoms in order to obtain ionization energies for different positions in
the cluster.
Furthermore, statistical effects were found to increase the decay width
for both the ICD and the ETMD. Since the ICD decay width in the model of pairs,
considering nearest neighbours only, scales linearly with the number
of nearest neighbours $N$ and the ETMD3 decay width in the model of triples
scales like $N(N-1)$,
the ETMD3 is statistically preferred and can therefore compete
with the usually faster ICD.

Based on the strong structure dependence of the secondary electron spectrum
observed during the PhD, a new
structure analysis method for noble gas clusters was developed. If two
competing ICD-like processes are energetically accessible and can be measured
independently, the comparison of the experimentally obtained cluster
composition and relative peak intensities can be compared to theoretically
modelled spectra for a large variety of structures. From the best agreement,
the mean cluster structure can be deduced as was carried out for
a set of experimentally created NeAr cluster manifolds. They were shown
to have a shell structure with an argon core of different sizes surrounded
by complete and incomplete shells of neon atoms.

In the future, it would be worth investigating \ac{ICD} processes with electron
transitions forbidden in the non-relativistic description.
Additionally, for the ETMD3 the determination of a physically reasonable
choice of the energy transfer distance $R$ would be beneficial.
Furthermore, the ETMD3 should be investigated using the relativistic
FanoADC-Stieltjes approach studying the scalar-relativistic effects observed
for the case of the Auger effect following an ionization of the Xe4d.
An analytic investigation of the angular distribution of the emitted ICD
electron would give further insights and could be compared to the experimental
data available.

It might be worth implementing the FanoADC based on a partitioning by energy in
order to obtain partial decay widths for the different terms stemming from the
same 2h configuration of the final states and to
improve their accuracy by including interchannel couplings. This approach
would also automatically allow for the treatment of systems, where the
atoms ionized in the final state can be generated by a symmetry operation
of the system's point group.
Furthermore, a more rigorous investigation of the couplings between the
initial state and Rydberg states would be beneficial for the development
of a method to exclude these couplings from the decay width calculation
automatically.
For decay width calculations for different interatomic distances,
a rapid change is observed at distances of about \unit[10 -- 15]{\AA}.
This behaviour is known for the non-relativistic treatment for all
investigated systems as well,
but the reason is so far unknown. It might be caused by insufficient
basis sets and additional ghost centers between the atoms of the decay partners
might shift this effect to larger distances. However, the reason for
this behaviour should be investigated and the results should be taken
into account in later studies.

In the Stieltjes Imaging, the pseudo-spectra are currently shifted such,
that the lowest energy has some predefined, small number. This number needs
to be small in order to allow for numerical stability for as high order
moments as possible. However, whether there is an optimal choice of this
number and if it exists, what it depends on, is currently unknown.
It would be worth investigating, because an improvement of the Stieltjes
imaging procedure might lead in more reliable results or decay width profiles,
which are easier to interprete, which might allow for a more automatized
evaluation of the results.
So far, no automatic error estimation is available in the FanoADC-Stieltjes
approach. However, it would be very helpful for the validation of the
calculated decay widths.
Further, the Stieltjes-Imaging applied to a combination of two or more
pseudo-spectra of the same system calculated using different basis sets
as shown in Ref. \cite{Reinhardt79} would be worth investigating.
In this way, the density of the pseudo-spectrum would ideally increase in the
energy region of interest. Therefore, a more accurate description of the
decay widths might be possible. At the same time, this approach might enable
decay width calculations of larger systems, where a huge basis set cannot be
applied due to lacking computational resources required in the integral
transformation step, which is the bottleneck of the calculations.

For the estimation of secondary electron spectra of clusters,
the investigation of cluster dynamics and their influence on the
secondary electron spectra should be developed and included in
HARDRoC. This knowledge about the oscillations and their displacements
would at the same time allow for an error estimation of
the calculated decay widths.

The investigated structures have so far been chosen by educated guesses.
For a more detailed study, optimized cluster structures would be beneficial.
Furthermore, it would be a challenging task to develop the possibility to
treat clusters with constituents of lower than spherical symmetry, e.g.,
water molecules.
