\chapter[Heteroatomic Noble Gas Clusters]{Competing Processes in Heteroatomic Noble Gas Clusters}
\label{chapter_clusters}

The modelling of secondary electron spectra originating from \ac{ICD}
and \ac{ETMD}3 processes are carried out using the program
HARDRoC \cite{HARDRoC}. It is based on the model of pairs and
triples explained in chapter \ref{chapter:geom} and is applied
to a given cluster structure. The 
decay widths obtained are proportional to the probability of the underlying decay.
The experimental spectra are obtained from a multitude of distinct measurements.
Hence, for a reasonable large number of measurements, a distribution with the
same statistical behaviour as for the decay width calculation in the
model of pairs and triples is achieved. Therefore, the experimental spectra
can be directly compared with the theoretically obtained spectra.

The energies of the secondary electrons are calculated using
equation \ref{equation:E_sec}. The single ionization energies
can be obtained in two different ways. They can be
estimated by the atomic ionization energies in Table
\ref{table:noble_atom_properties} corrected by energetic shifts
given in Table \ref{table:cluster_shifts} in order to take
the effect of the cluster environment into account. 
Alternatively, their source can be the single ionization spectra measured
at the same time and the
same conditions as the electron-electron
coincidence experiment. The latter apporach is to be preferred,
since the energy shifts
of Table \ref{table:cluster_shifts} are obtained from analysis of
homonuclear cluster's ionization spectra and
hence the values only give a first approximation to the real energetic shift in
heteronuclear cluster. Since both, the \ac{ICD} and primarily the
\ac{ETMD}3 occur at interfaces, it is crucial for a good modelling to
describe the ionization energies of these atoms as accurate as possible.

The decay widths $\Gamma$ can either be obtained from the asymptotic expressions
in equations (\ref{reltheolifetime_exp}) and (\ref{reltheolifetimeetmd_exp})
using the atomic properties given in Table \ref{table:noble_atom_properties}
or from a fit to decay widths obtained from FanoADC-Stieltjes calculations.

In the following sections two different heteronuclear clusters are investigated
illustrating different effects on the secondary electron spectra.
In the case of ArXe clusters, the focus is set on the investigation of how
spin-orbit coupling affects
the secondary electron spectra. Furthermore, I will investigate the influence
of the cluster size and number of argon shells around the xenon core and
the effects of of the basic structure of the cluster and discuss the
difference between the spectra
obtained from clusters having an icosahedral
and \ac{fcc} structure.
In the case of NeAr clusters I will focus on the bidirectional dependence of
the ICD spectra of the competing NeNe-ICD and NeAr-ICD processes and
different arrangements of neon atoms around the argon core.

\newpage
\input{subsecs/arxe_clusters}
\newpage
\input{subsecs/near_clusters}

